Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-27 analysis/paper/paper.Rmd
conclusion and a paragraph in background section

# Introduction and Research Question

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005]. Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale. Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005). Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004;@Mullins2011;@Silliman2001;@Rubertone2000;@Torrence2000;@TorrenceandClarke2000;@Voss2005).

@@

<--Add a paragraph about exchange network before European contact-->

@@

Indirect influence of colonialism has been proposed recently and some students in many parts of the world [@Trabert2017]. Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts.

>>>>

# Introduction

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005]. Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale. Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005). Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004;@Mullins2011;@Silliman2001;@Rubertone2000;@Torrence2000;@TorrenceandClarke2000;@Voss2005).

@@

Although the northeastern Taiwan forms a circumscribed environment due to the surrounding mountains and sea, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups in northern and southern Taiwan, foreign traders such as the Han people of China, and Europeans. This local exchange network of northeastern Taiwan had global connections after the 16th century because of its location on shipping routes between China and Japan (Chen 2005; Wang and Liu 2007). The groups involved in the exchange network consisted by the Chinese traders, indigenous groups called Basay in northern Taiwan, and indigenous groups, Kavalan, in northeastern Taiwan. Historical records indicate that the Kavalan people were one of the Basay’s trading partners, and they exchanged rice, deer skin, and gold for metal tools or ornaments, beads, and Chinese ceramics (Chen 2005; Hsieh 2009a; Li and Wu 2006).

@@

Indirect influence of colonialism has been proposed recently and some studies reveal how peripheral areas had been influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017]. Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts. Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper. As a status marker, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods. Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context. Those ornaments may convey different meaning with exotic and colonizer image that increases more competitions between aggrandized individuals for prestige and wealth accumulation, which might result in a increasing social inequality. The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern. This study demonstrate that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization. It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system. During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population. This is a case study that focuses on trade ornaments and more studies can be done to have a comparison with other archaeological evidence. Further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
